### 2015-06-28 reports/paper/experiment.tex
thesis update

By comparing experiments at normal speed (fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn}) with experiments at fast speed (fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf}), we find that the localization error does not increase when the movement speed increased from $10$ cm per second to $20$ cm per second. However, it's conceivable that the localization error will increase with further speed increase, but since our system is designed for HCI, we feel $20$ cm per second covers most of the use cases.

@@

\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{color_5.png}
    \caption{Painting an eclipse. The color of the heart changes with the volume of the song.}
    \label{fig:show_color_3}
\end{figure}

>>>>

By comparing experiments at normal speed (fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn}) with experiments at fast speed (fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf}), we find that the localization error does not increase when the movement speed increased from $10$ cm per second to $20$ cm per second. However, it's conceivable that the localization error will increase with further speed increase, but since our system is designed for HCI, a speed of $20$ cm per second covers most of the use cases.

@@

\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{color_5.png}
    \caption{Painting an ellipse. The color of the painting is controlled by the volume of the white noise.}
    \label{fig:show_color_3}
\end{figure}

### 2015-06-28 reports/paper/introduction.tex
thesis update

Global Positioning System (GPS) is the prevailing technology used for outdoor localization. A set of 32 satellites equipped with synchronized atomic clock orbit around the earth and broadcast their time and position information at fixed intervals. A GPS receiver listens to at least 4 satellites and uses the broadcasted timing information to infer the location. Commercial grade GPS has an average error of a few meters, depending on the size and quality of the receiver~\cite{intro:gps}. While accuracy in this range is good for many applications including driving navigation and vehicle tracking, it does not provide enough precision for local movement tracking.

Ultrasound based indoor localization approaches, on the other hand, have achieved sub-centimeter accuracy~\cite{intro:ultra}. These systems employ multiple ultrasound receivers and use the arrival time difference between the sound source and receivers to infer the source location. Although obtaining good accuracy, ultrasound systems require the use of expensive transducers.

Bluetooth and Wi-Fi based technologies have gained popularity in indoor positioning recently, mainly due to the widespread deployment of bluetooth tags and Wi-Fi stations in public spaces. The fact that modern consumer devices such as mobile phones and tablets are commonly equipped with Wi-Fi and bluetooth modules also made this approach particular attractive because no hardware needs to be installed on the user end. Relevant commercial user case includes applications that serve advertisements and coupons to consumers in shopping centers based on the customer's location. In these systems, device signal strength received at different base stations are used for the estimation of the user device location. However, their reported accuracy are in the range of $1$ to $5$ meters~\cite{intro:blue, intro:loc}, which is not accurate enough for local movement tracking.

>>>>

Global Positioning System (GPS) is the prevailing technology used for outdoor localization. A set of $32$ satellites equipped with synchronized atomic clock orbit around the earth and broadcast their time and position information at fixed intervals~\cite{intro:gps2}. A GPS receiver listens to at least 4 satellites and uses the broadcasted timing information to infer the location. Commercial grade GPS has an average error of a few meters, depending on the size and quality of the receiver~\cite{intro:gps}. While accuracy in this range is good for many applications including driving navigation and vehicle tracking, it does not provide enough precision for local movement tracking.

Ultrasound based indoor localization approaches, on the other hand, have achieved sub-centimeter accuracy~\cite{intro:ultra, intro:ultra2}. These systems employ multiple ultrasound receivers and use the arrival time difference between the sound source and receivers to infer the source location. Although obtaining good accuracy, ultrasound systems require the use of expensive transducers.

Bluetooth and Wi-Fi based technologies have gained popularity in indoor positioning recently, mainly due to the widespread deployment of bluetooth tags and Wi-Fi stations in public spaces. The fact that modern consumer devices such as mobile phones and tablets are commonly equipped with Wi-Fi and bluetooth modules also made this approach particular attractive because no hardware needs to be installed on the user end. Relevant commercial user case includes applications that serve advertisements and coupons to consumers in shopping centers based on the customer's location. In these systems, device signal strength received at different base stations are used for the estimation of the user device location. However, their reported accuracy are in the range of $1$ to $5$ meters~\cite{intro:blue, intro:loc, intro:wifi}, which is not accurate enough for local movement tracking.

### 2015-06-28 reports/paper/prior.tex
thesis update

\subsection{ILD}
ILD techniques rely on the observation that signal intensity decays as the distance to the microphone increases. A microphone closer to the signal source would receive the signal with higher intensity than a microphone farther away. With multiple microphones, it is possible to infer the source location by comparing the signal intensity received at different microphones. Human's auditory system has used ILD cues to infer source direction~\cite{ild:2}, and this technique is most effective when used to localize high frequency sources, because these sources don't diffract significantly around the listener's head and produce a significant intensity difference.

@@

The second benefit of formulating cross correlation in frequency domain is that it provides a unified framework to prefilter the signals. Cross-correlation with prefiltering is known as \emph{generalized cross correlation (GCC)}. Different prefiltering approaches have been investigated to improve arrival time difference estimation~\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}. Under the GCC framework, the arrival time difference $t_0$ between two signals $x_1(t)$ and $x_2(t)$ is estimated as:
\begin{eqnarray} \label{eq:gcc}
t_0 &=& \arg\max_{\tau} R_{x_1x_2}(\tau) \\\label{eq:gcc2}
R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
\end{eqnarray}
$W(\omega)$ provides a way to prefilter signals passed to the cross correlation estimator. We focused on three ways of prefiltering the signal:
\begin{description}%[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
\item[GCC] $W(\omega) = 1$. No prefiltering is done. This is unfiltered normal cross correlation.
\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divided by its magnitude. Only phase information contributes to delay estimation.
\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. Part of magnitude information is included in arrival time difference estimation.
\end{description}

>>>>

\subsection{ILD}
ILD techniques rely on the observation that signal intensity decays as the distance to the microphone increases. A microphone closer to the signal source would receive the signal with higher intensity than a microphone farther away. With multiple microphones, it is possible to infer the source location by comparing the signal intensity received at different microphones. Human's auditory system has used ILD cues to infer source direction~\cite{ild:2, ild:h, ild:5}, and this technique is most effective when used to localize high frequency sources, because these sources don't diffract significantly around the listener's head and produce a significant intensity difference.

@@

The second benefit of formulating cross correlation in frequency domain is that it provides a unified framework to prefilter the signals. Cross-correlation with prefiltering is known as \emph{generalized cross correlation (GCC)}. Different prefiltering approaches have been investigated to improve arrival time difference estimation~\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}. Under the GCC framework, the arrival time difference $t_0$ between two signals $x_1(t)$ and $x_2(t)$ is estimated as:
\begin{eqnarray} \label{eq:gcc}
t_0 &=& \arg\max_{\tau} R_{x_1x_2}(\tau) \\\label{eq:gcc2}
R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
\end{eqnarray}
$W(\omega)$ provides a way to prefilter signals passed to the cross correlation estimator. We focused on three ways of prefiltering the signal:
\begin{description}%[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
\item[GCC] $W(\omega) = 1$. No prefiltering is done. This is unfiltered normal cross correlation.
\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divided by its magnitude. Only phase information contributes to delay estimation.
\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. Part of the magnitude information is included in arrival time difference estimation.
\end{description}

### 2015-06-28 reports/paper/system.tex
thesis update

\subsection{Hardware}
The end system has two arrays, each with three microphones mounted on the vertices of a $20$ cm equilateral triangle. We used omni-directional foil electret condenser microphone due to its low cost and small size. The microphone has a frequency range of $100$ to $10K$ Hz, and a minimum SNR of 58dB\cite{sys:1}. We also used operational amplifer OPA344 from Texas Instrument to amplify the microphone output by a factor of 100, so the received signal can be easily picked up by the analog-to-digital converter (ADC) module installed on the micro-controller. The micro-controller board used in this project is \emph{teensy 3.1}, and it is attached to one of the vertices of the triangle. Fig~\ref{fig:setup_array} shows a picture of the array setup.  The micro-controller contains an onboard RAM of $64$k, and an ADC module capable of sampling at $500$kHz~\cite{tdoa:micloc, sys:teensy}. In this project, the micro-controller collects microphone data on all three channels for a duration $12$ milliseconds and then sends the recorded data to a computer through the USB port for localization.

@@

On the software side, \emph{Python} is used as the main programming language since it has extensive libraries in both real time data handling and signal processing. \emph{ZeroMQ} is an inter process messaging queue that is used in our system to pass data across different modules. Our system is made up of two data acquisition modules (each is used to receive raw date from microphone array output) and one localization module (listens to both data acquisition modules and perform localization using the raw microphone data). Using ZeroMQ as connections to different modules adds flexibility to our system, as we can design applications such as the drawing application we have used in our system to interface only with the localization module and disregard how the microphone data is collected. Other localization applications can be easily integrated into our system by connecting them with the localization module. Furthermore, we have built a recording module that interfaces with the data acquisition modules for offline data analysis and parameter tuning.

>>>>

\subsection{Hardware}
The end system has two arrays, each with three microphones mounted on the vertices of a $20$ cm equilateral triangle. We used omni-directional foil electret condenser microphone due to its low cost and small size. The microphone has a frequency range of $100$ to $10K$ Hz, and a minimum SNR of 58dB\cite{sys:1}. We also used operational amplifer OPA344 from Texas Instrument to amplify the microphone output by a factor of 100, so the received signal can be easily picked up by the analog-to-digital converter (ADC) module installed on the micro-controller. The micro-controller board used in this project is \emph{teensy 3.1}, and it is attached to one of the vertices of the triangle. Fig~\ref{fig:setup_array} shows a picture of the array setup.  The micro-controller contains an onboard RAM of $64$k, and an ADC module capable of sampling at $500$kHz~\cite{tdoa:micloc, sys:teensy, sys:teensy2}. In this project, the micro-controller collects microphone data on all three channels for a duration $12$ milliseconds and then sends the recorded data to a computer through the USB port for localization.

@@

On the software side, \emph{Python} is used as the main programming language since it has extensive libraries in both real time data handling and signal processing. \emph{ZeroMQ} is an inter process messaging queue that is used in our system to pass data across different modules. Our system is made up of two data acquisition modules (each is used to receive raw data from microphone array output) and one localization module (listens to both data acquisition modules and perform localization using the raw microphone data). Using ZeroMQ as connections to different modules adds flexibility to our system, as we can design applications such as the drawing application we have used in our system to interface only with the localization module and disregard how the microphone data is collected. Other localization applications can be easily integrated into our system by connecting them with the localization module. Furthermore, we have built a recording module that interfaces with the data acquisition modules for offline data analysis and parameter tuning.
